---
title: "Covid91 vaccine study Final2024"
author: "Amir Johnson"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### BarPlot
```{r}
Males <-filter(FinalDat, sex=="M")
```


```{r}
Males <-select(Males,-sex)
```

```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This analysis includes two plots illustrating the relationship between COVID-19 infection rates and the administration of either a placebo or vaccine among the male population.

The first plot shows that 26% of individuals who received the vaccine developed COVID-19, whereas around 74% of those who received the placebo contracted the disease. This indicates that the vaccine significantly reduces the likelihood of contracting COVID-19 compared to the placebo, despite some vaccine recipients still becoming infected. For those who received the placebo, there is a 3:1 chance of contracting COVID-19. Men who took the medication had a 1.5-fold lower risk of infection compared to those who took the placebo.

The two plots are described as follows:
1. In the first plot, the x-axis represents infection status (whether participants have COVID-19), and the y-axis shows the count of individuals who either did or did not contract the disease.
2. The second plot focuses on those with COVID-19, depicted by the two bars on the left. The x-axis shows 
COVID-19 status, while the y-axis displays the percentage of patients who contracted COVID-19 after receiving
either the vaccine or the placebo, rather than the total number of cases.In summary, the first plot 
illustrates infection counts, while the second plot shows the percentage of cases among vaccinated versus 
placebo groups.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

According to the data, 74% of males who received a placebo and 26% who received the vaccine contracted COVID-19. This indicates that men are less likely to contract COVID-19 when they receive the vaccine compared to the placebo. The vaccine proves to be a more effective option than the placebo, as receiving the placebo results in twice the likelihood of contracting COVID-19.

### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```


The results from Fisher's Exact test should be carefully considered in light of these findings. The p-value reported is 2.2e-16, which is exceptionally small. This tiny value suggests that the observed outcomes are highly unlikely to have occurred by chance.

The null hypothesis posits that sex does not influence COVID-19 mortality. However, the evidence strongly supports the alternative hypothesis, which suggests a significant relationship between sex and COVID-19 death rates.



```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```

```{r}
low=(Malesfish$conf.int[1]/(Malesfish$conf.int[1] +1))*100
hi=(Malesfish$conf.int[2]/(Malesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The efficacy of the COVID-19 vaccine for men is seen above. The vaccine is effective if the efficacy rate is higher than 50%. We can conclude that the vaccine is unsuccessful in preventing men from contracting COVID-19 based on the interval shown above. 

The approximate effectiveness of the vaccine is indicated by the odds ratio. As can be seen above, the men' odds ratio is 74. For the males, the effectiveness confidence interval is hi 78.32 to low 68,84. 

## Females
### Barplot



```{r}
Females <-filter(FinalDat, sex=="F")
```


```{r}
Females <-select(Females,-sex)
```



```{r}
dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot examines the female population's COVID-19 infection rates in relation to whether they received the vaccine or a placebo.

Compared to the previous plot, this data shows a higher incidence of COVID-19 among those who received the placebo. Specifically, only about 35% of individuals who received the vaccine contracted COVID-19, while approximately 65% of those who got the placebo ended up with the disease. This suggests that receiving the vaccine provides better protection against COVID-19 than the placebo, with a 1:2 chance of infection for those vaccinated. Girls who received the vaccine had a 50% chance of contracting COVID-19 compared to those who received the placebo.

The analysis includes two plots The first plot shows the x-axis as infection status, indicating whether participants have COVID-19, and the y-axis represents the count of individuals who either did or did not contract the disease.


### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The percentages of girls, or women, who contracted COVID-19 after receiving the vaccine and a placebo are displayed in these numerical results. 

35% of the gals (women) who had the vaccination and developed COVID-19 were infected, as roughly shown by the earlier plots. 65% of the girls who received the placebo also had COVID-19. Girls are indeed effectively shielded from contracting COVID-19 by the immunization.

For girls receiving the vaccine, the risk of contracting COVID-19 is 1:1. The risk of contracting COVID-19 was halved for the girls who received the vaccine compared to the placebo. group.
## LGBTQ



## Druggies


# Overall Results and Conclusions
