---
title: "Covid91 vaccine study Final2024"
author: "Amir Johnson"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### BarPlot
```{r}
Males <-filter(FinalDat, sex=="M")
```


```{r}
Males <-select(Males,-sex)
```

```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This analysis includes two plots illustrating the relationship between COVID-19 infection rates and the administration of either a placebo or vaccine among the male population.

The first plot shows that 26% of individuals who received the vaccine developed COVID-19, whereas around 74% of those who received the placebo contracted the disease. This indicates that the vaccine significantly reduces the likelihood of contracting COVID-19 compared to the placebo, despite some vaccine recipients still becoming infected. For those who received the placebo, there is a 3:1 chance of contracting COVID-19. Men who took the medication had a 1.5-fold lower risk of infection compared to those who took the placebo.

The two plots are described as follows:
1. In the first plot, the x-axis represents infection status (whether participants have COVID-19), and the y-axis shows the count of individuals who either did or did not contract the disease.
2. The second plot focuses on those with COVID-19, depicted by the two bars on the left. The x-axis shows 
COVID-19 status, while the y-axis displays the percentage of patients who contracted COVID-19 after receiving
either the vaccine or the placebo, rather than the total number of cases.In summary, the first plot 
illustrates infection counts, while the second plot shows the percentage of cases among vaccinated versus 
placebo groups.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

According to the data, 74% of males who received a placebo and 26% who received the vaccine contracted COVID-19. This indicates that men are less likely to contract COVID-19 when they receive the vaccine compared to the placebo. The vaccine proves to be a more effective option than the placebo, as receiving the placebo results in twice the likelihood of contracting COVID-19.

### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```


The results from Fisher's Exact test should be carefully considered in light of these findings. The p-value reported is 2.2e-16, which is exceptionally small. This tiny value suggests that the observed outcomes are highly unlikely to have occurred by chance.

The null hypothesis posits that sex does not influence COVID-19 mortality. However, the evidence strongly supports the alternative hypothesis, which suggests a significant relationship between sex and COVID-19 death rates.



```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```

```{r}
low=(Malesfish$conf.int[1]/(Malesfish$conf.int[1] +1))*100
hi=(Malesfish$conf.int[2]/(Malesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The efficacy of the COVID-19 vaccine for men is seen above. The vaccine is effective if the efficacy rate is higher than 50%. We can conclude that the vaccine is unsuccessful in preventing men from contracting COVID-19 based on the interval shown above. 

The approximate effectiveness of the vaccine is indicated by the odds ratio. As can be seen above, the men' odds ratio is 74. For the males, the effectiveness confidence interval is hi 78.32 to low 68,84. 

## Females
### Barplot



```{r}
Females <-filter(FinalDat, sex=="F")
```


```{r}
Females <-select(Females,-sex)
```



```{r}
dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot examines the female population's COVID-19 infection rates in relation to whether they received the vaccine or a placebo.

Compared to the previous plot, this data shows a higher incidence of COVID-19 among those who received the placebo. Specifically, only about 35% of individuals who received the vaccine contracted COVID-19, while approximately 65% of those who got the placebo ended up with the disease. This suggests that receiving the vaccine provides better protection against COVID-19 than the placebo, with a 1:2 chance of infection for those vaccinated. Girls who received the vaccine had a 50% chance of contracting COVID-19 compared to those who received the placebo.

The analysis includes two plots The first plot shows the x-axis as infection status, indicating whether participants have COVID-19, and the y-axis represents the count of individuals who either did or did not contract the disease.


### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The percentages of girls, or women, who contracted COVID-19 after receiving the vaccine and a placebo are displayed in these numerical results. 

35% of the gals (women) who had the vaccination and developed COVID-19 were infected, as roughly shown by the earlier plots. 65% of the girls who received the placebo also had COVID-19. Girls are indeed effectively shielded from contracting COVID-19 by the immunization.

For girls receiving the vaccine, the risk of contracting COVID-19 is 1:1. The risk of contracting COVID-19 was halved for the girls who received the vaccine compared to the placebo. group.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Femalesfish <- fisher.test(table1)
Femalesfish
```

The output of the Fisher's Exact test should be carefully examined among all of these findings. At 0.95, the p-value 2e-08 shown above is incredibly tiny. This is a minuscule amount. This number indicates that the results are unlikely to be the result of chance, and the alternative hypothesis can be accepted in place of the null hypothesis.


According to the Fishers' Exact Test, the odds ratio was 1.9. Women who received the medication had an approximate 0.32-fold increased risk of contracting COVID-19 compared to those who received a placebo



```{r}
effect <- Femalesfish$p.value/(Femalesfish$p.value +1)
effectivenessFemales = Femalesfish$estimate/(Femalesfish$estimate +1)*100
print(effectivenessFemales)
```

```{r}
low=(Femalesfish$conf.int[1]/(Femalesfish$conf.int[1] +1))*100
hi=(Femalesfish$conf.int[2]/(Femalesfish$conf.int[2] +1))*100
options(digits=2)
```



## Effectiveness Confidence Interval `r low`  to `r hi`


The COVID-19 effectiveness for gals (females) is shown above. The vaccine is effective if the efficacy rate is higher than 65%. We can conclude that the vaccination is very successful in preventing girls from contracting COVID-19 based on the interval shown above. 

70% or higher is the standard for a vaccine's effectiveness. This vaccination is quite effective at preventing women from contracting COVID-19.



## LGBTQ

### Barplot

```{r}
LGBTQ <-filter(FinalDat, LGBTQ=="gay")
```


```{r}
dd2 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The chart above illustrates the relationship between LGBTQ individuals and their likelihood of contracting COVID-19 based on whether they received a vaccination or a placebo.

The bar plot reveals that 19% of LGBTQ individuals received the vaccine while 80% received the placebo. Those who received the placebo had a higher probability of contracting COVID-19 compared to those who were vaccinated.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data reveal the percentages of LGBTQ individuals who contracted COVID-19 after receiving either the vaccine or a placebo.

The results show that 20% of LGBTQ participants who received the vaccine contracted COVID-19, while 80% of those who received the placebo did. This indicates that the placebo offers no effective protection against COVID-19 for the LGBTQ population and is not recommended.

There is a 3.2:1 likelihood of contracting COVID-19 when receiving the placebo. Individuals who received the vaccine and identified as LGBTQ were three times less likely to contract COVID-19 compared to those who received the placebo.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```

The p-value displayed above is extremely small, at 2e-06. This is a really little amount. This number suggests that these results are not likely to have occurred by coincidence.

The null hypothesis states that there is no relationship between sex and COVID-19-related mortality. The overwhelming evidence that sex does affect people's death rates points to the alternative theory, which holds that there is a link between LGBTQ and COVID-19 deaths.



```{r}
effect <- LGBTQfish$p.value/(LGBTQfish$p.value +1)
effectivenessLGBTQ = LGBTQfish$estimate/(LGBTQfish$estimate +1)*100
print(effectivenessLGBTQ)
```

```{r}
low=(LGBTQfish$conf.int[1]/(LGBTQfish$conf.int[1] +1))*100
hi=(LGBTQfish$conf.int[2]/(LGBTQfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The data above shows the COVID-19 vaccine's effectiveness for LGBTQ(gay). If the vaccination's efficacy rate is more than 50%, it is considered effective. Based on the time frame above, we may conclude that the vaccine is not effective in protecting LGBTQ(gay) from acquiring COVID-19. 

The odds ratio shows how effective the vaccination is, roughly. The odds ratio for LGBTQ(gay) is 81, as can be shown above. The effectiveness confidence interval for men is hi 89.87 to low 68.99.



## Druggies

# Barplot

```{r}
Drugs <-filter(FinalDat, DrugUser=="no")
```

```{r}
Drugs <-select(Drugs,-DrugUser)
```


```{r}
dd2 <-  Drugs %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This figure illustrates the connection between drug use and the likelihood of contracting COVID-19 based on whether individuals received the vaccine or a placebo.

The bar plot shows that approximately 51% of respondents who reported drug use and received the vaccine contracted COVID-19, while nearly 49% of drug users who received the placebo also contracted the virus. This indicates a 1:1 chance of contracting COVID-19 with the vaccine. However, the risk of contracting COVID-19 was halved for drug users who received the vaccine compared to those who received the placebo.

There are two plots The first plot has the x-axis representing infection status (whether participants have COVID-19), and the y-axis shows the count of individuals who contracted the disease versus those who did not.The second plot features the left two bars representing all individuals with COVID-19. The x-axis indicates COVID-19 status, while the y-axis displays the percentage of patients who contracted the disease after receiving either the vaccine or placebo, rather than the total number of cases.

# Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Drugs)
rowPerc(table1)
colPerc(table1)
```

The percentages of drug users who contracted COVID-19 after receiving the vaccine and a placebo are displayed in these numerical results.
Thirty-three percent of drug users who received the vaccination also contacted COVID-19. About 67% of patients who received merely a placebo also had COVID-19. It is evident that the COVID-19 vaccine works to protect drug users from COVID-19, hence receiving this vaccination is advised. There is a 1:2 chance of contracting COVID-19 while receiving the vaccination. The risk of contracting COVID-19 was halved among drug users who received the vaccine compared to those who received a placebo.

# Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
Drugsfish <- fisher.test(table1)
Drugsfish
```

There is a 0 p-value. Given the tiny to none quantity, it may be concluded that these results are unlikely to be the result of chance.

The results of the Fishers' Exact Test show that the odds ratio was 3, meaning that drug users who received the vaccine had a roughly 40-50% lower chance of contracting Covid-19 than those who received a placebo.

The alternative is accepted and the null hypothesis is rejected. The medication works well to lower drug users' risk of contracting COVID-19.

```{r}
effect <- Drugsfish$p.value/(Drugsfish$p.value +1)
effectivenessDrugs = Drugsfish$estimate/(Drugsfish$estimate +1)*100
print(effectivenessDrugs)
```

```{r}
low=(Drugsfish$conf.int[1]/(Drugsfish$conf.int[1] +1))*100
hi=(Drugsfish$conf.int[2]/(Drugsfish$conf.int[2] +1))*100
options(digits=2)
```

# Effectiveness Confidence Interval `r low`  to `r hi`

The COVID-19 effectiveness for men is seen above. The vaccine is effective if the efficacy rate is higher than 40%. We can conclude that the vaccine effectively prevents drug users from contracting COVID-19 based on the period shown above. 

60% or higher is the standard for a vaccine's effectiveness. When it comes to preventing drug users from getting COVID-19, this vaccine is very effective.

The approximate effectiveness of the vaccine is indicated by the odds ratio. The odds ratio for drug users is 75, as can be shown above. For drug users, the efficacy confidence interval ranges from 71.42 to 78.22.


# Overall Results and Conclusions
