---
title: "Covid91 vaccine study Final2024"
author: "Amir Johnson"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
Males <-filter(FinalDat, sex=="M")
```


```{r}
Males <-select(Males,-sex)
```

```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This analysis includes two plots illustrating the relationship between COVID-19 infection rates and the administration of either a placebo or vaccine among the male population.

The first plot shows that 26% of individuals who received the vaccine developed COVID-19, whereas around 74% of those who received the placebo contracted the disease. This indicates that the vaccine significantly reduces the likelihood of contracting COVID-19 compared to the placebo, despite some vaccine recipients still becoming infected. For those who received the placebo, there is a 3:1 chance of contracting COVID-19. Men who took the medication had a 1.5-fold lower risk of infection compared to those who took the placebo.

The two plots are described as follows:
1. In the first plot, the x-axis represents infection status (whether participants have COVID-19), and the y-axis shows the count of individuals who either did or did not contract the disease.
2. The second plot focuses on those with COVID-19, depicted by the two bars on the left. The x-axis shows 
COVID-19 status, while the y-axis displays the percentage of patients who contracted COVID-19 after receiving
either the vaccine or the placebo, rather than the total number of cases.In summary, the first plot 
illustrates infection counts, while the second plot shows the percentage of cases among vaccinated versus 
placebo groups.

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions
